---
title: "Datapack COP 20 Code to Combine and Give us an Output"
author: "EA Team"
date: "Last Updated:4/17/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---
**Instructions for recreating ER 'dashboard dataset' using COP20 Datapack and FAST tool for countries that aren't part of the regional programs.

Step 1: Install the required packages to your computer. This includes loading the SI tools from Github (be sure to set up a Github beforehand). You should only have to do this part in the initial set up. Once doing that, skip to the next section during recurring downloads. 

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
install.packages("devtools")
install.packages("tidyverse")
devtools::install_github("USAID-OHA-SI/tameDP")
devtools::install_github("USAID-OHA-SI/fastR")
library(stringr)
library(dplyr)
library(tidyverse)
library(readxl)
library(ICPIutilities)
library(glamr)
library(here)
library(fastR)
library(tameDP)
library(purrr)
here("Datasets")
```

```
#Identify all the FAST files(Where they are located on your computer)
```{r,results="Hide"}
fasts <- list.files(path="FASTs", full.names = TRUE)
```

#Create lists of all datapacks
```{r,results="Hide"}
datapacks <- list.files("Data Pack", full.names = TRUE)
```
#step 3: filpath to the dashboard dataset (excel sheet). It best to move this to where your WD is
```{r,results="Hide"}
dataset <- "Budget.ER.MER Clean Post FY 19 1.18.20v2 (1).xlsx"
```

#Read in all FAST files and combine into one data frame. This pulls in the intervention tab
```{r,results="Hide"}
df_fasts <- map_dfr(.x = fasts,
                  .f = ~ fastR::run_fastR_cop20(.x, "2 Intervention-E"))
```
#Data Cleaning. This includes renaming, and filtering out certain values
```{r,results="Hide"}
df_fasts <- df_fasts %>%
  dplyr::rename(`Funding Agency` = fundingagency,
                `Mechanism Name` = mechanismname,
                `Partner Name` = primepartner,
                `Partner Org Type` = orgtype,
                `Mechanism ID` = mech_code,
                `Mechanism Name` = mechanismname,
                `Program Area` = program,
                `Sub-Program Area` = programarea,
                `Service Delivery` = servicedelivery,
                `Sub-Beneficiary` = subbeneficiary,
                Beneficiary = beneficiary,
                `OU (COP 2019)` = ou,
                Budget = amt) %>%
  dplyr::filter(cop == "COP20",
                amt_type == "Budget")

df_fasts <- df_fasts %>%
  dplyr::mutate(`Fiscal Year` = "FY 21",
                `Data Stream` = "Budget.PA",
                `Local Partner Type` = "Unknown",
                Country = `OU (COP 2019)`,
                `OU (Pre-COP 2019)` = `OU (COP 2019)`) %>%
  dplyr::select(-cop, -amt_type)
```

#read in all of the datapacks and combine into one data frame
```{r,results="Hide"}

df_dp <- map_dfr(.x = datapacks,
                   .f = ~ tameDP::tame_dp(.x))
 
df_dp <- df_dp %>%
  dplyr::filter(indicator %in% c("HTS_TST", "HTS_TST_POS", "TX_CURR", "TX_NEW", "VMMC_CIRC", "OVC_SERV", "PrEP_CURR", "PrEP_NEW"),
                disaggregate != "KeyPop/Result") %>%
  dplyr::group_by(operatingunit, fundingagency, mech_code, primepartner, mech_name, indicator, fiscal_year) %>%
  dplyr::summarise(targets = sum(targets, na.rm = TRUE)) %>% 
  dplyr::ungroup()

df_dp <- df_dp %>%
  dplyr::rename(`OU (COP 2019)` = operatingunit,
                `Funding Agency` = fundingagency,
                `Mechanism ID` = mech_code,
                `Partner Name` = primepartner,
                `Mechanism Name` = mech_name,
                `Fiscal Year` = fiscal_year,
                Indicator = indicator,
                Target = targets) %>%
  dplyr::mutate(`OU (Pre-COP 2019)` = `OU (COP 2019)`,
                `Target Result` = "Target",
                `Fiscal Year` = "FY 21",
                `Data Stream` = "MER",
                Country = `OU (COP 2019)`) %>%
  dplyr::mutate(`Program Area` = dplyr::case_when(Indicator == "HTS_TST_POS" ~ "HTS",
                                                  Indicator == "HTS_TST" ~ "HTS",
                                                  Indicator == "TX_CURR" ~ "C&T",
                                                  Indicator == "TX_NEW" ~ "C&T",
                                                  Indicator == "PrEP_CURR" ~ "PREV",
                                                  Indicator == "PrEP_NEW" ~ "PREV",
                                                  Indicator == "VMMC" ~ "PREV"))

```
## read in and clean up dashboard dataset and creates the integrated data pack
``````{r,results="Hide"}
data <- readxl::read_excel(dataset) %>%
  dplyr::mutate(`Mechanism ID` = as.character(`Mechanism ID`),
                Budget = as.numeric(Budget),
                Target = as.numeric(Target),
                Result = as.numeric(Result))
```
## now bind them together and can reorder columns if you want (for EATAP work)
```{r,results="Hide"}

#col_order<-c ("OU (COP 2019)", "OU (Pre-COP 2019)","Country","Funding Agency","Partner Name","Mechanism Name","Mechanism ID","Data Stream","Fiscal Year","Budget","Program Area","Sub-Program Area","Service Delivery","Beneficiary","Sub-Beneficiary","Partner Org Type","Local Partner Type","int#","isindigenouspartner")
#df<-df_fasts[, col_order]
#df$"Program Area: Sub-Program Area-Service Level"<-paste(df$`Program Area`,": ",df$`Sub-Program Area`,"-",df$`Service Delivery`)
#df$"Beneficiary-Sub-Beneficiary"<-paste(df$Beneficiary,"-",df$`Sub-Beneficiary`)


df <- dplyr::bind_rows(df_fasts, df_dp,)
```
#Print statements (Print(df) or View(df)) can be a helfpul way to check if your data is transforming the way you want it to
``````{r,results="Hide"}
df <- dplyr::bind_rows(df_fasts, df_dp, data)
```
##write file to output folder
```{r,results="Hide"}
write_csv(df, "Budget.ER.MER.dd.mm.yyyy.csv")

df_dp %>% distinct(indicator, disaggregate) %>% prinf
```
