---
title: "EA Integrated Financial Dataset"
author: "EA team"
date: "September 24, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Please be sure to install these if you do not have the following packages on your computer
library(stringr)
library(dplyr)
library(tidyverse)
library(readxl)
library(ICPIutilities)
library(glamr)
library(purrr)
library(googlesheets4)
library(tidylog) #used to track merges
library(here)
here("Datasets")
```



Below is a guide for working with the EA branches FSD_Clean and MSD_Clean functions. You can use these functions to importat, clean, and append the Financial Structured Dataset and MER Structured Dataset to create an integrated financial and programmatic dataset. 

1. Use this to read in and clean the MSD
```{r} 
MSD_Clean<-function(df){
#Can run these separately if need be. read_MSD is nested here
  df<-read_msd(df)

#Use this to filter out the disaggs and drop uneeded columns
  df<-df%>%dplyr::filter(standardizeddisaggregate == "Total Numerator") %>%
    dplyr::select( - c('operatingunituid', 'pre_rgnlztn_hq_mech_code', 'prime_partner_duns',	'award_number',	
'categoryoptioncomboname', 'ageasentered', 'trendsfine', 'trendssemifine',	
'statushiv',	'statustb', 'statuscx', 'hiv_treatment_status', 'otherdisaggregate', 'otherdisaggregate_sub', 'source_name','numeratordenom','disaggregate','standardizeddisaggregate','trendscoarse','sex','modality'))%>%
    dplyr::rename("Operating Unit"= "operatingunit",
                  "Country"= "countryname",
                  "Funding Agency"= "fundingagency",
                  "Prime Partner Name"= "primepartner",
                  "Mechanism ID"="mech_code",
                  "Mechanism Name" = "mech_name",
                  "Indicator Type"="indicatortype",
                  "Indicator"="indicator",
                  "Targets"="targets",
                  "Quarter 1"="qtr1",
                  "Quarter 2"="qtr2",
                  "Quarter 3"="qtr3",
                  "Quarter 4"="qtr4",
                  "Cumulative"="cumulative",
                  "Fiscal Year"="fiscal_year")%>%
    dplyr::mutate(`Data Stream`="MER")%>%
      dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`))

#Turn NAs into 0s to make it easier to read
  df<-df%>%
    mutate_at(vars(`Quarter 1`:`Quarter 4`),~replace_na(.,0))%>% 
    mutate_at(vars(`Targets`),~replace_na(.,0))%>%
    mutate_at(vars(`Cumulative`),~replace_na(.,0))
    
    
#using pivot_long to turn results into quarters
df<-df%>%
    pivot_longer(
      cols = `Quarter 1`: `Quarter 4`,
      names_to="Quarter",
      values_to="Results"
    )
   
  return(df)
}
```

2. clean the FSD function
```{r} 
FSD_Clean<-function(df){

#nested read_msd. Can be removed and run separately
  df<-read_msd(df)

# Drop columns you don't need and rename
  df<-df %>%dplyr::select( - c('prime_partner_duns',	'award_number',	
                                       'subrecipient_duns')) %>% 
    dplyr::rename("Operating Unit"= operatingunit,
                  "Country"= country,
                  "Funding Agency"= fundingagency,
                  "Prime Partner Name"= prime_partner_name,
                  "Subrecipient Name"= subrecipient_name,
                  "Mechanism ID"=mech_code,
                  "Mechanism Name" = mech_name,
                  "Program Area"= program,
                  "Sub Program Area" = sub_program,
                  "Interaction Type"= interaction_type,
                  "Beneficiary" = beneficiary,
                  "Sub Beneficiary"= sub_beneficiary,
                  "Cost Category"= cost_category,
                  "Sub Cost Category" =sub_cost_category,
                  "Fiscal Year" = fiscal_year,
                  "COP Budget New Funding"=cop_budget_new_funding,
                  "COP Budget Pipeline"=cop_budget_pipeline,
                  "Total Planned Funding" = cop_budget_total,
                  "Workplan Budget" = workplan_budget_amt,
                  "Expenditure"=expenditure_amt,
                  "Prime Partner Type"=prime_partner_org_type,
                  "Is Indigenous Prime Partner"=is_indigenous_prime_partner)%>%
    dplyr::mutate(`Data Stream`="FSD")

#replace NAs with 0s
    df<-df%>%
      mutate_at(vars(`COP Budget New Funding`:`Expenditure`),~replace_na(.,0))

    #add in a quarter for the budget data, this is useful for doing quarterly analytics of budget data
    df$Quarter<-c("Quarter 1")

#convert budget columns to numeric
    df<-df%>%
      dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`))%>%
      dplyr::mutate(`COP Budget New Funding`=as.numeric(`COP Budget New Funding`))%>%
      dplyr::mutate(`COP Budget Pipeline`=as.numeric(`COP Budget Pipeline`))%>%
      dplyr::mutate(`Total Planned Funding`=as.numeric(`Total Planned Funding`))%>%
      dplyr::mutate(`Expenditure`=as.numeric(`Expenditure`))
    

  return(df)
}
```
3. Read in the MSD/FSD and clean it using our functions. Pattern is helpful to use here but be careful that you are not picking up more than one dataset
```{r}
#Here you are pointing the script to your "Datasets" folder to search for that pattern
filesMSD<-list.files("Datasets",pattern="MER_Structured_Datasets_OU_IM_FY18-21_20200918_v2_1",full.names = TRUE)
df.MSD<-purrr::map_dfr(.x=filesMSD,
                    .f=~MSD_Clean(.x))
filesFSD<-list.files("Datasets",pattern="Financial_Structured_Dataset_FY18-21_20200918_v2_1",full.names = TRUE)
df.FSD<-purrr::map_dfr(.x=filesFSD,
                       .f=~FSD_Clean(.x))
```

4. Bind the ouput (if applicable)
```{r} 
df.FSDMSD<-bind_rows(df.FSD,df.MSD)%>%

#add in OU column
  dplyr::mutate(`OU` = `Operating Unit`)%>%
  
#Add in agency category column to group agencies
  dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
  mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                    ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                    ifelse(`Agency Category` =="Dedup", "Dedup","Other"))))
  
#re-order the columns
df.FSDMSD<-df.FSDMSD%>%
  dplyr::relocate(`OU`,`Operating Unit`:`Fiscal Year`,`Data Stream`,`COP Budget New Funding`:`Expenditure`)
 
```
Read in the historical base dataset for FY15-17. It is kept in the EA team google folder
``` {r}
#BudH<-read_sheet("https://docs.google.com/spreadsheets/d/1Y_u4smwYUgSKSirbihRZ8Rh8GE1yA31-NlSg5GnzHsU/edit#gid=1536348418")

BudH<-read_xlsx(here("Datasets","Historic Budget ER MER fy15-17.xlsx"))
 BudH<-BudH%>%
      dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`))%>%
      dplyr::mutate(`Mechanism ID`=as.character(`Mechanism ID`))%>%
      dplyr::mutate(`Cost Category`=as.character(`Cost Category`))%>%
      dplyr::mutate(`Sub Cost Category`=as.character(`Sub Cost Category`))%>%
      dplyr::mutate(`Total Planned Funding`=as.numeric(`Total Planned Funding`))%>%
      dplyr::mutate(`Expenditure`=as.numeric(`Expenditure`))%>%
      dplyr::mutate(`Targets`=as.double(`Targets`))%>%
      dplyr::mutate(`Results`=as.double(`Results`))
 BudH<-BudH%>%
   mutate_at(vars(`Expenditure`:`Total Planned Funding`, `Targets`:`Results`),~replace_na(.,0))
 
 #drop the oldpartner types column as this will lead to duplicates if you wish to update it
 BudH<-BudH%>%
   dplyr::select(-c(`Partner Type-USAID Internal`, `InitiativeName`:`Funding Account`))
   

df.EA<-bind_rows(df.FSDMSD,BudH)
df.EA<-df.EA%>%
  dplyr::relocate(`OU`,`Operating Unit`:`Fiscal Year`,`Data Stream`,`COP Budget New Funding`:`Expenditure`)

# Update Site Level
df.EA<-df.EA%>%
  dplyr::mutate(`Site Level` = `Program Area`)
  df.EA<-df.EA%>%
    dplyr::mutate(`Site Level` = ifelse(`Site Level` == "C&T" , "Site",
                                  ifelse(`Site Level` == "PREV" , "Site",
                                  ifelse(`Site Level` == "HTS" , "Site",
                                  ifelse(`Site Level` == "SE" , "Site",
                                    ifelse(`Site Level` == "ASP", "Above Site",
                                    ifelse(`Site Level` == "AP", "AP",
                                    ifelse(`Site Level` =="PM", "PM","N/A"))))))))
#df.EA<-df.EA%>%
  #dplyr::mutate(`Interaction Type1`=`Interaction Type`)
#df.EA<-df.EA%>%
  #dplyr::mutate(`Interaction Type1`= ifelse(`Program Area` == "PM", "PM",.))
```
Update the USAID Internal partners list. I downloaded it from the LP team. The end goal will be to have a better integration between LP and the "is indigenous" column. This is a temporary solution hopefully
```{r}
Partner_Types<-read_xlsx(here("Datasets","Partner types 9.24.2020.xlsx"))
Partner_Types<-Partner_Types%>%
  dplyr::mutate(`Mechanism ID`= as.character(`Mechanism ID`))
df.EA<-left_join(df.EA,Partner_Types,by ="Mechanism ID", name="Partner Types-USAID Internal")

  
```

Fun extra piece if you'd like to triangulate the data with outlays
```{r}
filesOutlay<-list.files("Datasets",pattern="outlay meeting",full.names = TRUE)
outlay<-readxl::read_xlsx(filesOutlay)
outlay<-outlay%>%
  dplyr::mutate_at(vars(`Outlay`:`Planning Level`),~replace_na(.,0))%>%
  dplyr::mutate(`Outlay` = as.integer(`Outlay`))%>%
  dplyr::mutate(`Planning Level` = as.integer(`Planning Level`))%>%
  dplyr::mutate(`Fiscal Year` = as.character(`Fiscal Year`))%>%
  dplyr::rename(`Prime Partner Name` = "Partner Name")

EA.Integrated.Budget.MER.Outlay<-bind_rows(outlay,df.EA)%>%
  left_join(EA.Integrated.Budget.MER.Outlay,Partner_Types,by="Mechanism ID")%>%
   dplyr::filter(`Funding Agency`=="USAID")%>%
  dplyr::filter(`Fiscal Year`=="2020")

```

5. Write CSV
```{r echo=T, results='hide'} 
write_csv(df.EA, "EA Integrated Financial Dataset.10.30.2020.csv")

```
6. Send to google drive (beta). This takes a long time and isn't recommended
```{r}

ss1 <- gs4_create(
  "Budgter.MER.FSD. 9.22.2020",
  sheets = "chickwts")%>%
sheet_write(df.FSDMSD, ss = ss1, sheet = "chickwts")

```



  